Near and Mid-infrared properties of known $z\geq5$ Quasars

### 2019-06-12 tex/VHzQ.tex
tex/VHzQ update; light_curves/lightcurve_plot.py udpate

Using the extended datasets described in Section~\ref{sec:NIR_data}, 
we select objects with at least 8 measurements observed with intervals of 
at least 30 days, in at least one filter. 
The the average calibrated flux over all epochs must be $>0$ 
(i.e. {\tt aperJky3}$>0$) 
with the signal-to-noise is required to be $\geq8$. 
(i.e. {\tt aperJky3/aperJky3Err}$>=$8).
The clipped median and standard deviation is then calculated 
\begin{equation}
      var = 1.48 \times  {\rm m.a.d.} / \bar{\epsilon}
\end{equation}
where $var$ is the index of variation, m.a.d. is the median absolute
deivation, and $\bar{\epsilon}$ is the mean of the error in each point
in the light curve, divided by the total number of points.  With these
criteria, there are 21 WFCAM and 12 VISTA objects with 1 object (VIK
J1148+0056) in common to both, and this sample of 32 is our starting
point for variability investigations.

We present four objects that are particularly well-sampled in the NIR. 
These are: 
MMTJ0215-0529 (Figure~\ref{fig:MMTJ0215-0529}), 
CFHQSJ0216-0455 (Figure~\ref{fig:CFHQSJ0216-0455}), 
SHELLQsJ0220-0432 (Figure~\ref{fig:SHELLQsJ0220-0432}) and 
SDSS J0959+0227 (Figure~\ref{fig:SDSSJ0959+0227}). 
We incorporate the MIR NEOWISE-R light curve data where available.

\begin{figure}
  \begin{subfigure}{}
    \centering
    \includegraphics[width=8.5cm]{../light_curves/MMTJ0215-0529LC_20190214.png}
    \caption{The infrared light curve for MMT J0215-0529 (}
    \label{fig:MMTJ0215-0529}
  \end{subfigure} 
  \begin{subfigure}{}
    \centering
    \includegraphics[width=8.5cm]{../light_curves/CFHQSJ0216-0455LC_20190214.png}
    \caption{CFHQS J0216-0455}
    \label{fig:CFHQSJ0216-0455}
  \end{subfigure}
\end{figure}

\begin{figure}
  \begin{subfigure}{}\quad
    \centering
    \includegraphics[width=8.5cm]{../light_curves/SHELLQsJ0220-0432LC_20190214.png}
    \caption{SHELLQs J0220-0432}
    \label{fig:SHELLQsJ0220-0432}
  \end{subfigure}
  \begin{subfigure}{}\quad
    \centering
    \includegraphics[width=8.5cm]{../light_curves/SDSSJ0959+0227LC_20190612.png}
    \caption{SDSS J0959+0227}
    \label{fig:SDSSJ0959+0227}
  \end{subfigure}
  \medskip
\end{figure}

Even with well sampled data across the 3000-4000 day observed time-scales that the WFCAM and VISTA surveys span, the $(1+z)$ time-dilation dramatically affects the sampled rest-frame timescales sampled.

SDSS J0959+0227 is in the VH$z$Q sample from \citet{Wang2016} 
%% Wang et al. 2016, ApJ, 819, 24, and their Table1. 
but is first reported in \citet{Civano2011} as CID-2220 (CXOCJ095908.1+022707), 
%2011, ApJ, 741, 91 
a high-redshift, $z>3$ AGN in the Chandra-COSMOS field. 
%%
The spectrum from this object is presented in \citet{Ikeda2012} 
% et al., 2012, ApJ, 756, 160, 
and is called object ``No. 11'' in that papers' Table 1. 
and it is a beauty!! Figure 2 with the spectrum is attached. 
\citet{Ikeda2012} show SDSS J0959+0227 having a narrow Ly$\alpha$ line, 
and would likely be a Lyman-$\alpha$ emitting galaxy had it not been an X-ray source, 
The X-ray luminosity is $\approx3x10^44$ erg s$-1$ in the 2–10 keV rest-frame, 
so it is an AGN. This object is clearly not a regular broadline ``Type 1'' AGN, but  
if there is even the slightest of variability, this could potentially be a high-$z$ 
AGN transitioning from a narrow-line ``Type 2'' object to a broadline Type 1 quasar.

Doing this independently in each band...

>>>>

Using the extended datasets described in Section~\ref{sec:NIR_data},
we select objects with at least 8 measurements observed with intervals
of at least 30 days, in at least one filter. The the average
calibrated flux over all epochs must be $>0$ (i.e. {\tt aperJky3}$>0$)
with the signal-to-noise is required to be $\geq8$. (i.e. {\tt
aperJky3/aperJky3Err}$\geq$8). The clipped median and standard
deviation is then calculated
\begin{equation}
      var = 1.48 \times  {\rm m.a.d.} / \bar{\epsilon}
\end{equation}
where $var$ is the index of variation, m.a.d. is the median absolute
deivation, and $\bar{\epsilon}$ is the mean of the error in each point
in the light curve, divided by the total number of points.  With these
criteria, there are 21 WFCAM and 12 VISTA objects with 1 object (VIK
J1148+0056) in common to both, and this sample of 32 is our starting
point for variability investigations.

We present four objects that are particularly well-sampled in the NIR. 
These are: 
MMT J0215-0529       ($z=5.13$; Figure~\ref{fig:MMTJ0215-0529}), 
CFHQS J0216-0455    ($z=6.01$; Figure~\ref{fig:CFHQSJ0216-0455}), 
SHELLQs J0220-0432 ($z=5.90$; Figure~\ref{fig:SHELLQsJ0220-0432}) 
and 
SDSS J0959+0227 ($z=5.07$; Figure~\ref{fig:SDSSJ0959+0227}). 
We incorporate the MIR NEOWISE-R light curve data where available.

\begin{figure}
  \begin{subfigure}{}
    \centering
    \includegraphics[width=8.5cm]{../light_curves/MMTJ0215-0529LC_20190214.png}
    \caption{The infrared light curve for MMT J0215-0529 with data from 
      WFCAM (smaller solid cirlces), VIRCAM (hexagons) and NEOWISE-R (larger solid cirlces).}
    \label{fig:MMTJ0215-0529}
  \end{subfigure} 
  \begin{subfigure}{}
    \centering
    \includegraphics[width=8.5cm]{../light_curves/CFHQSJ0216-0455LC_20190214.png}
    \caption{The same as Fig.~\ref{fig:MMTJ0215-0529} except for CFHQS J0216-0455.}
    \label{fig:CFHQSJ0216-0455}
  \end{subfigure}
\end{figure}

\begin{figure}
  \begin{subfigure}{}\quad
    \centering
    \includegraphics[width=8.5cm]{../light_curves/SHELLQsJ0220-0432LC_20190214.png}
    \caption{The same as Fig.~\ref{fig:MMTJ0215-0529} except for SHELLQs J0220-0432.}
    \label{fig:SHELLQsJ0220-0432}
  \end{subfigure}
  \begin{subfigure}{}\quad
    \centering
    \includegraphics[width=8.5cm]{../light_curves/SDSSJ0959+0227LC_20190612.png}
    \caption{The near infrared light curve for SDSS J0959+0227. Note, 
      SDSS J0959+0227 is too faint in the MIR to be detected by ALLWISE or unnWISE.}
    \label{fig:SDSSJ0959+0227}
  \end{subfigure}
  \medskip
\end{figure}

Even with well sampled data across the 3000-4000 day observed
time-scales that the WFCAM and VISTA surveys span, the $(1+z)$
time-dilation dramatically affects the sampled rest-frame timescales
sampled. Indeed, when sharp changes in a accretion rate are expected
on the system's dynamical timescale of several kiloyears
\citep[e.g.,][]{Regan2019}, then seeing any variability signature is
not expected. However, noting how in lower redshift quasars, which
also have massive $>10^{8}$ M$_{\odot}$ black holes, dramatic changes
in both continuum and line emission {\it is seen} on much shorter
timescales, continued monitoring of these objects is warrented.

We finally focus on SDSS J0959+0227 which is in the VH$z$Q sample from
\citet{WangF2016} but is first reported in \citet{Civano2011} as
CID-2220 and a high-redshift, $z>3$ AGN, in the Chandra-COSMOS
field. The spectrum from this object is presented in \citet{Ikeda2012}
and shows SDSS J0959+0227 having a narrow Ly$\alpha$ line, and would
likely be a Lyman-$\alpha$ emitting galaxy had it not been an X-ray
source, The X-ray luminosity is $\approx3x10^44$ erg s$-1$ in the 2–10
keV rest-frame, so it is an AGN. This object is clearly not a regular
broadline ``Type 1'' AGN, but if with the slightest hint of
variability, this could potentially be a high-$z$ AGN transitioning
from a narrow-line ``Type 2'' object to a broadline Type 1 quasar.
